### 2018-11-08 R/transect-analysis.Rmd
Organize data, start analysis notebook.

---
title: "Upper Ouémé Forest Fungi -- Soil Transect Analysis"
author: "Brendan Furneaux"
date: '`r paste0("October 10, 2018; Last modified ", strftime(file.mtime(knitr::current_input()), "%B %d, %Y"))`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries.
```{r}
library(magrittr)
library(tidyverse)
library(readxl)
```

Define the required files.
```{r}
base.dir <- str_extract(getwd(), ".+oueme-fungi-transect")
data.dir <- file.path(base.dir, "data")
soils.dir <- file.path(data.dir, "soils")
soils.file <- file.path(soils.dir, "transectsoil_corrected.csv")

lab.dir <- file.path(data.dir, "lab_setup")
plate.file <- file.path(lab.dir, "Brendan soil.xlsx")

pacbio.dir <- file.path(data.dir, "PacBio")
pacbio.otu.file <- file.path(pacbio.dir, "OTU_table.txt")

iontorrent.dir <- file.path(data.dir, "IonTorrent")
iontorrent.otu.file <- file.path(iontorrent.dir, "OTU_,matrix.txt")
```

Load the soil chemistry data
```{r}
soils <- read_tsv(soils.file, locale = locale(decimal_mark = ",")) %>%
  extract("Echantillon", into = c("Plot", "Sample", "Date"),
          regex = "(V1P[123])T([0-9]+)-([0-9]{4}-[0-9]{1,2}-[0-9]{1,2})") %>%
  mutate_at("Date", as.Date)
```

Load the data for the sample placements in the plates
```{r}
plate <- read_xlsx(plate.file, "Concentration samples", skip = 1, col_names = TRUE) %>%
  rename(Comment = X__1, Well = row) %>%
  select (-X__2)

```

Load the IonTorrent OTU table
```{r}
ion.otu <- read_tsv(iontorrent.otu.file) %>%
  column_to_rownames("OTU") %>%
  as.matrix
```

```{r}
pb.otu <- read_tsv(pacbio.otu.file)

pb.otumatrix <- pb.otu %>%
  select(Group:H7) %>%
  column_to_rownames("Group") %>%
  as.matrix
```