### 2018-12-18 R/transect-analysis.Rmd
Working on new demultiplexing and dada2 clustering, use Makefile

---
title: "Upper Ouémé Forest Fungi -- Soil Transect Analysis"
author: "Brendan Furneaux"
date: '`r paste0("October 10, 2018; Last modified ", strftime(file.mtime(knitr::current_input()), "%B %d, %Y"))`'
output:
  beamer_presentation:
    keep_tex: TRUE
  bookdown::pdf_document2:
    keep_tex: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 5, fig.height = 5)
library(magrittr)
library(tidyverse)
library(readxl)
library(rwantshue)
colorscheme <- iwanthue(seed=12, force_init = TRUE)
doctype <- knitr::opts_knit$get("rmarkdown.pandoc.to")

if (! is.null(doctype) && doctype == "beamer") {
  knitr::opts_chunk$set(message = FALSE, warning = FALSE, out.width = "100%", fig.height = 4, fig.width = 8)
}
```

# Doctype

Doctype is `r doctype`.

```{r include = FALSE}
base.dir <- str_extract(getwd(), ".+oueme-fungi-transect")
data.dir <- file.path(base.dir, "data")
soils.dir <- file.path(data.dir, "soils")
soils.file <- file.path(soils.dir, "transectsoil_corrected.csv")

lab.dir <- file.path(data.dir, "lab_setup")
plate.file <- file.path(lab.dir, "Brendan_soil.xlsx")
tags.file <- file.path(lab.dir, "Hectors_tag_primer_plates.xlsx")

pacbio.dir <- file.path(data.dir, "PacBio")
pacbio.otu.file <- file.path(pacbio.dir, "PACBIO_OTU_table.xlsx")
pacbio.funguild.file <- file.path(pacbio.dir, "PacBio_FUNGuild_in.guilds.web.txt")
pacbio.group.files <- list.files(pacbio.dir, "*.groups")

iontorrent.dir <- file.path(data.dir, "IonTorrent")
iontorrent.otu.file <- file.path(iontorrent.dir, "higherLvl", "OTU.txt")
iontorrent.classification.file <- file.path(iontorrent.dir, "hiera_BLAST.txt")
iontorrent.funguild.file <- file.path(iontorrent.dir, "IonTorrent_FUNGuild.guilds.txt")
iontorrent.repseq.file <- file.path(iontorrent.dir, "Otus_rep_seqs.fa")
```

```{r}
# Load the soil chemistry data
soils <- read_tsv(soils.file, locale = locale(decimal_mark = ",")) %>%
  extract("Echantillon", into = c("Plot", "Sample", "Date"),
          regex = "(V1P[123])T([0-9]+)-([0-9]{4}-[0-9]{1,2}-[0-9]{1,2})") %>%
  mutate_at("Date", as.Date)
```

```{r}
# Load the data for the sample placements in the plates

true_tags <- read_xlsx(plate.file, "Taggar fITS7", skip = 1) %>%
  map(list(1:3, 5:7, 9:11), ~ select(.y, !!.x), .) %>%
  map_dfr(set_names, c("Well", "No", "Tag"))

wrong_tags <- read_xlsx(tags.file, skip = 1) %>%
  select(Well = LOCATION, oligoname) %>%
  mutate(Tag = str_replace(oligoname, "gITS7mod-", "") %>% as.numeric)

well_key <- wrong_tags %>%
  left_join(select(true_tags, Tag, True_Well = Well))

plate <- read_xlsx(plate.file, "Concentration samples", skip = 1, col_names = TRUE) %>%
  rename(Comment = X__1, Well = row) %>%
  select (-X__2) %>%
  # Sample number consists of site, location along transect, and possible qualifier
  tidyr::extract("Sample", c("Site", "X", "Qual"), "(V1P[13]|T[12])[ -][ST](\\d+)(\\D*)$") %>%
  mutate_at("X", as.integer) %>%
  # Sites were recorded differently in the two years
  mutate(Site = case_when(Site == "V1P1" ~ "Ang",
                          Site == "V1P3" ~ "Gan",
                          Site == "T1" ~ "Gan",
                          Site == "T2" ~ "Ang",
                          TRUE ~ NA_character_)) %>%
  # For some site/years, coordinates were measured from the edge of the plot rather than numbered from 1--25 
  group_by(Year, Site) %>%
  mutate(X = X - if (!any(is.na(X)) && max(X) > 25) 12 else 0) %>%
  ungroup %>%
  tidyr::extract("Well", c("Row", "Column"), "([A-H])(\\d+)", remove = FALSE) %>%
  mutate_at("Row", factor, levels = LETTERS) %>%
  mutate_at(c("Column", "Row"), as.integer) %>%
  mutate(sample_type = factor(Comment, levels = c("Blank", "Pos", "Pos. Kontroll")) %>%
           fct_collapse(Blank = "Blank", Pos = c("Pos", "Pos. Kontroll")) %>%
           fct_explicit_na("Sample"))

plate %>% group_by(Year, Site) %>%
  summarize_at("X", c("min", "max"))
```

```{r}
# Load the IonTorrent OTU table
ion.otu <- read_tsv(iontorrent.otu.file) %>%
  mutate(n_reads = rowSums(select_if(., is_numeric))) %>%
  left_join(read_tsv(iontorrent.classification.file)) %>%
  select(-matches("[A-H]\\d+"), -contains("ITS")) %>%
  left_join(read_tsv(iontorrent.funguild.file) %>% rename(OTU = Group)) %>%
  select(-matches("[A-H]\\d+"), -contains("ITS")) %>%
  rename(Kingdom = Domain) %>%
  dplyr::mutate_at(dplyr::vars(Phylum:Species), dplyr::na_if, "?") %>%
  dplyr::mutate_at(dplyr::vars(Phylum:Species), dplyr::na_if, "unidentified") %>%

  # make "unident." as informative as possible
  dplyr::mutate_at("Kingdom", replace_na, "unident.") %>%
  dplyr::mutate(Phylum = ifelse(is.na(Phylum),
                                ifelse(is.na(Kingdom),
                                       NA_character_,
                                       paste0("unident. ", Kingdom)),
                                Phylum),
                Class = ifelse(is.na(Class),
                                ifelse(is.na(Phylum),
                                       NA_character_,
                                       paste0("unident. ", Phylum)),
                                Class),
                Order = ifelse(is.na(Order),
                                ifelse(is.na(Class),
                                       NA_character_,
                                       paste0("unident. ", Class)),
                                Order),
                Family = ifelse(is.na(Family),
                                ifelse(is.na(Order),
                                       NA_character_,
                                       paste0("unident. ", Order)),
                                Family),
                Genus = ifelse(is.na(Genus),
                                ifelse(is.na(Family),
                                       NA_character_,
                                       paste0("unident. ", Family)),
                                Genus),
                Species = ifelse(is.na(Species),
                                ifelse(is.na(Genus),
                                       NA_character_,
                                       paste0(Genus, " sp.")),
                                Species)) %>%
  mutate_at(dplyr::vars(Kingdom:Species),
            str_replace,
            "(unident\\. )*unident(ified|\\.)",
            "ZZZZunident.") %>%

  # Sort taxonomically. Put the "unident." at the end
  arrange(Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
  mutate_at(dplyr::vars(Kingdom:Species),
            str_replace,
            "ZZZZ",
            "") %>%

  # Add the representative sequence for each OTU and calculate the length
  left_join(
    seqinr::read.fasta(file = iontorrent.repseq.file, as.string = TRUE) %>%
    { tibble(OTU = names(.),
             RepSeq = unlist(.),
             SeqLen = str_length(RepSeq)) }) %>%

  # Remove the positive control
  filter(is.na(Species) | Species != "Agaricus bisporus") %>%

  # Keep the data in this order.
  mutate_at(dplyr::vars(Kingdom:Species),
            ~ factor(.) %>% fct_inorder)

ion.platemap <- read_tsv(iontorrent.otu.file) %>%
  column_to_rownames("OTU") %>%
  as.matrix %>%
  colSums %>%
  tibble(Well = names(.), n_reads = .) %>%
  mutate_at("Well", plyr::mapvalues, well_key$Well, well_key$True_Well, FALSE) %>%
  full_join(plate %>%
              select(Plate, Well, sample_type) %>%
              full_join(
                expand.grid(Plate = 1:2,
                            Row = LETTERS[1:8],
                            Column = 1:12) %>%
                  tidyr::unite(Well, Row, Column, sep = ""))) %>%
  mutate(Plate = "IonTorrent") %>%
  mutate_at("sample_type", fct_explicit_na, "None") %>%
  group_by_if(~ !is.factor(.)) %>%
  summarize(sample_type = paste(as.character(sort(unique(sample_type))), collapse = "/")) %>%
  ungroup %>%
  mutate(f_reads = n_reads / max(n_reads)) %>%
  tidyr::extract("Well", c("Row", "Column"), "([A-H])(\\d+)") %>%
  mutate_at("Row", factor, levels = LETTERS) %>%
  mutate_at(c("Column", "Row"), as.integer)
```

# Demultiplexing

\tiny

We used a forward primer which included the forward IonTorrent sequencing adapter and an index, which distinguishes the well the sample was place in in the plate.

\center

```{tikz out.width = "60%"}
\begin{tikzpicture}[line width=7]

\newcommand{\fadapt}{\draw[red] (start) -- node[midway, below]{fwd adapter}  ++(3, 0) coordinate (start);}
\newcommand{\well}{\draw[blue] (start) -- node[midway, above]{well index}  ++(1, 0) coordinate (start);}
\newcommand{\gITS}{\draw[black!30!green] (start) -- node[midway, below]{gITS7}  ++(1.5, 0) coordinate (start);}
\newcommand{\seq}{\draw[black] (start) -- ++(1, 0) coordinate (start);
\draw[black, dotted] (start) -- node[midway, below]{sequence} ++(2,0) coordinate(start);
\draw[black] (start) -- ++(1,0) coordinate (start);}
\newcommand{\ITS}{\draw[cyan] (start) -- node[midway, below]{ITS4}  ++(1.5, 0) coordinate (start);}
\newcommand{\plate}{\draw[orange!40!yellow] (start) -- node[midway, above]{plate index}  ++(1, 0) coordinate (start);}

\newcommand{\radapt}{\draw[pink] (start) -- node[midway, below]{rev adapter}  ++(3, 0) coordinate (start);}

\newcommand{\fprimer}{\coordinate (pstart) at (start);
\fadapt
\well
\gITS
\coordinate (end) at (start);
\path (end) ++(0,0.75) coordinate (end);
\draw[line width = 2, dashed, |-|] (pstart) ++(0, 0.75) -- node[midway, above]{fwd primer}  (end);}

\newcommand{\rprimer}{\coordinate (pstart) at (start);
\ITS
\coordinate (end) at (start);
\path (end) ++(0,0.751) coordinate (end);
\draw[line width = 2, dashed, |-|] (pstart) ++(0,0.75) -- node[midway, above]{rev primer}  (end);}

\coordinate (start) at (0,0);
\fprimer
\seq
\rprimer
\end{tikzpicture}
```

# Demultiplexing

\tiny

Because we had more than 96 samples, we used two plates.  The same indexes were used for both plates.

During IonTorrent library prep, sequencing adapters were ligated on to the amplicons.  The forward adapters included index barcodes to distinguish the two plates.  The orientation of the adapters was presumably random with respect to our amplicons.

\center

```{tikz out.width = "60%"}
\begin{tikzpicture}[line width=7]

\newcommand{\fadapt}{\draw[red] (start) -- node[midway, below]{fwd adapter}  ++(3, 0) coordinate (start);}
\newcommand{\well}{\draw[blue] (start) -- node[midway, above]{well index}  ++(1, 0) coordinate (start);}
\newcommand{\gITS}{\draw[black!30!green] (start) -- node[midway, below]{gITS7}  ++(1.5, 0) coordinate (start);}
\newcommand{\seq}{\draw[black] (start) -- ++(1, 0) coordinate (start);
\draw[black, dotted] (start) -- node[midway, below]{sequence} ++(2,0) coordinate(start);
\draw[black] (start) -- ++(1,0) coordinate (start);}
\newcommand{\ITS}{\draw[cyan] (start) -- node[midway, below]{ITS4}  ++(1.5, 0) coordinate (start);}
\newcommand{\plate}{\draw[orange!40!yellow] (start) -- node[midway, above]{plate index}  ++(1, 0) coordinate (start);}

\newcommand{\radapt}{\draw[pink] (start) -- node[midway, below]{rev adapter}  ++(3, 0) coordinate (start);}

\newcommand{\fprimer}{\coordinate (pstart) at (start);
\fadapt
\well
\gITS
\coordinate (end) at (start);
\path (end) ++(0,0.75) coordinate (end);
\draw[line width = 2, dashed, |-|] (pstart) ++(0, 0.75) -- node[midway, above]{fwd primer}  (end);}

\newcommand{\rprimer}{\coordinate (pstart) at (start);
\ITS
\coordinate (end) at (start);
\path (end) ++(0,0.751) coordinate (end);
\draw[line width = 2, dashed, |-|] (pstart) ++(0,0.75) -- node[midway, above]{rev primer}  (end);}

\coordinate (start) at (0,0);
\fprimer
\seq
\rprimer

\coordinate (start) at (-4, -4);
\fadapt\plate\fprimer\seq\rprimer\radapt

\coordinate (start) at (-3, -6.5);
\radapt\fprimer\seq\rprimer\plate\fadapt
\end{tikzpicture}
```

# Demultiplexing

\tiny

Prior to sequencing, IonTorrent uses surface-bound amplification to cover beads with many copies of the same sequence.  The forward adapter binds to the surface.

Our samples all contained two copies of the forward adapter.  Which one was used in each amplification step is random, but over the course of the entire amplification, the ones which result in shorter amplicons should be favored.

\center

```{tikz out.width = "60%"}
\begin{tikzpicture}[line width=7]

\newcommand{\fadapt}{\draw[red] (start) -- node[midway, below]{fwd adapter}  ++(3, 0) coordinate (start);}
\newcommand{\well}{\draw[blue] (start) -- node[midway, above]{well index}  ++(1, 0) coordinate (start);}
\newcommand{\gITS}{\draw[black!30!green] (start) -- node[midway, below]{gITS7}  ++(1.5, 0) coordinate (start);}
\newcommand{\seq}{\draw[black] (start) -- ++(1, 0) coordinate (start);
\draw[black, dotted] (start) -- node[midway, below]{sequence} ++(2,0) coordinate(start);
\draw[black] (start) -- ++(1,0) coordinate (start);}
\newcommand{\ITS}{\draw[cyan] (start) -- node[midway, below]{ITS4}  ++(1.5, 0) coordinate (start);}
\newcommand{\plate}{\draw[orange!40!yellow] (start) -- node[midway, above]{plate index}  ++(1, 0) coordinate (start);}

\newcommand{\radapt}{\draw[pink] (start) -- node[midway, below]{rev adapter}  ++(3, 0) coordinate (start);}

\newcommand{\fprimer}{\coordinate (pstart) at (start);
\fadapt
\well
\gITS
\coordinate (end) at (start);
\path (end) ++(0,0.75) coordinate (end);
\draw[line width = 2, dashed, |-|] (pstart) ++(0, 0.75) -- node[midway, above]{fwd primer}  (end);}

\newcommand{\rprimer}{\coordinate (pstart) at (start);
\ITS
\coordinate (end) at (start);
\path (end) ++(0,0.751) coordinate (end);
\draw[line width = 2, dashed, |-|] (pstart) ++(0,0.75) -- node[midway, above]{rev primer}  (end);}

\coordinate (start) at (0,0);
\fprimer
\seq
\rprimer

\coordinate (start) at (-4, -4);
\fadapt\plate\fprimer\seq\rprimer\radapt

\coordinate (start) at (-3, -6.5);
\radapt\fprimer\seq\rprimer\plate\fadapt

\coordinate (start) at (0, -10.5);
\fprimer\seq\rprimer\radapt

\coordinate (start) at (0, -13);
\fprimer\seq\rprimer\plate\fadapt
\end{tikzpicture}
```

# Plates
```{r platemap}
pacbio.group.files %>%
  file.path(pacbio.dir, .) %>%
  map_dfr(read_csv) %>%
  group_by(group) %>%
  summarize(n_reads = n()) %>%
  ungroup %>%
  tidyr::extract(group, c("Plate", "Well"), "(\\w+_\\d+)_([A-H]\\d+)") %>%
  tidyr::extract("Well", c("Row", "Column"), "([A-H])(\\d+)") %>%
  mutate_at("Row", factor, levels = LETTERS) %>%
  mutate_at(c("Column", "Row"), as.integer) %>%
  group_by(Plate) %>%
  mutate(f_reads = n_reads / max(n_reads)) %>%
  ggplot(aes(Column, Row, size = f_reads)) + 
  facet_wrap(facets = "Plate", ncol = 3) + 
  # geom_raster(aes(fill = sample_type)) +
  geom_point(aes(alpha = ifelse(f_reads > 0, 1, 0))) +
  # scale_fill_discrete(name = "Type") +
  scale_size(name = "Normalized\nReads", range = c(0.1, 3), trans = "log1p") +
  scale_alpha(guide = FALSE) +
  ggplot2::coord_equal() +
  scale_y_reverse(breaks = 1:8, labels = LETTERS[1:8]) +
  scale_x_continuous(breaks = 1:12, limits = c(0.5, 12.5)) +
  # scale_fill_brewer(name = "", type = "qual", palette = 3) +
  theme(legend.box = "vertical", legend.position = "bottom", legend.key.size = unit(4, "mm"))

pb.otu <- read_excel(pacbio.otu.file)

pb.otumatrix <- pb.otu %>%
  column_to_rownames("Group") %>%
  select(A7_483_1:H7_483_1) %>%
  as.matrix %>%
  t %>%
  as.tibble(rownames = "Sample") %>%
  tidyr::extract("Sample", c("Well", "Plate"), regex = "([A-H]1?[0-9])_483_([12])", remove = FALSE) %>%
  mutate_at("Plate", as.integer) %>%
  mutate_at("Well", plyr::mapvalues, well_key$Well, well_key$True_Well, FALSE)

pb.otumatrix %>%
  full_join(plate %>% select(Plate, Well, sample_type)) %>%
  tidyr::extract("Well", c("Row", "Column"), "([A-H])(\\d+)") %>%
  mutate_at("Row", factor, levels = LETTERS) %>%
  mutate_at(c("Column", "Row"), as.integer) %>%
  mutate_at("Plate", paste, "PacBio") %>%
  mutate(n_reads = rowSums(select(., starts_with("Otu")))) %>%
  full_join(expand(., Row, Column, Plate)) %>%
  mutate(n_reads = ifelse(is.na(n_reads), 0, n_reads)) %>%
  group_by(Plate) %>%
  mutate(f_reads = n_reads / max(n_reads),
         sample_type = ifelse(is.na(sample_type), "None", as.character(sample_type))) %>%
  ungroup %>%
  select(Row, Column, Plate, n_reads, f_reads, sample_type) %>%
  bind_rows(ion.platemap) %>%
  mutate_at("sample_type", factor, levels = c("Sample", "Pos", "Blank", "Pos/Sample", "Blank/Sample", "Sample/None", "Pos/None", "Blank/None", "None")) %>%

  ggplot(aes(Column, Row, size = f_reads)) + 
  facet_wrap(facets = "Plate", ncol = 3) + 
  geom_raster(aes(fill = sample_type)) +
  geom_point(aes(alpha = ifelse(f_reads > 0, 1, 0))) +
  scale_fill_discrete(name = "Type") +
  scale_size(name = "Normalized\nReads", range = c(0.1, 3), trans = "log1p") +
  scale_alpha(guide = FALSE) +
  ggplot2::coord_equal() +
  scale_y_reverse(breaks = 1:8, labels = LETTERS[1:8]) +
  scale_x_continuous(breaks = 1:12, limits = c(0.5, 12.5)) +
  scale_fill_brewer(name = "", type = "qual", palette = 3) +
  theme(legend.box = "vertical", legend.position = "bottom", legend.key.size = unit(4, "mm"))
```

```{r}
# Parse out the identification and classification
pb.otu %<>% select(-(A1_483_1:`ITS1-4_500_1`)) %>%
  #parse the name
  dplyr::bind_cols(
    stringr::str_split(.$X__2, stringr::fixed("|"), n = 5, simplify = TRUE) %>%
      tibble::as_data_frame() %>%
      set_names(c("Subject.Name", "Accession", "Species.Hypothesis",
                  "Ref.Type", "Classification"))) %>%
  tidyr::extract("Classification",
                 c("Kingdom", "Phylum", "Class", "Order",
                                 "Family", "Genus", "Species"),
                 "k__(.+)_p__(.+)_c__(.+)_o__(.+)_f__(.+)_g__(.+)_s__(.+)") %>%
  dplyr::select(-X__2) %>%

  # calculate the length of each sequence
  dplyr::mutate(SeqLen = str_length(RepSeq)) %>%

  #change "unident." to NA
  dplyr::mutate_at(dplyr::vars(Phylum:Species), dplyr::na_if, "unident.") %>%
  # make "unident." as informative as possible
  dplyr::mutate_at("Kingdom", replace_na, "unident.") %>%
  dplyr::mutate(Phylum = ifelse(is.na(Phylum),
                                ifelse(is.na(Kingdom),
                                       NA_character_,
                                       paste0("unident. ", Kingdom)),
                                Phylum),
                Class = ifelse(is.na(Class),
                                ifelse(is.na(Phylum),
                                       NA_character_,
                                       paste0("unident. ", Phylum)),
                                Class),
                Order = ifelse(is.na(Order),
                                ifelse(is.na(Class),
                                       NA_character_,
                                       paste0("unident. ", Class)),
                                Order),
                Family = ifelse(is.na(Family),
                                ifelse(is.na(Order),
                                       NA_character_,
                                       paste0("unident. ", Order)),
                                Family),
                Genus = ifelse(is.na(Genus),
                                ifelse(is.na(Family),
                                       NA_character_,
                                       paste0("unident. ", Family)),
                                Genus),
                Species = ifelse(is.na(Species),
                                ifelse(is.na(Genus),
                                       NA_character_,
                                       paste0(Genus, " sp.")),
                                Species)) %>%
  mutate_at(dplyr::vars(Kingdom:Species),
            str_replace,
            "(unident\\. )*unident(ified|\\.)",
            "ZZZZunident.") %>%

  # Sort taxonomically. Put the "unident." at the end
  arrange(Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
  mutate_at(dplyr::vars(Kingdom:Species),
            str_replace,
            "ZZZZ",
            "") %>%
  mutate_at(dplyr::vars(Kingdom:Species),
            ~ factor(.) %>% fct_inorder) %>%

  # Remove the positive control
  filter(is.na(Species) | Species != "Agaricus_bisporus") %>%

  # Add guild assignments
  left_join(read_tsv(pacbio.funguild.file))

```

```{r}
# Parse the total sequence count
pb.otu %<>%
  rename(n_reads = `OTU_ID=seqs_no`) %>%
  mutate_at("n_reads", str_replace, "Otu\\d+=", "") %>%
  mutate_at("n_reads", as.integer)
```

```{r}
all.otu <- bind_rows(
  mutate(pb.otu, tech = "PacBio"),
  mutate(ion.otu, tech = "IonTorrent")
) %>%
  mutate_at(dplyr::vars(Kingdom:Species),
            str_replace,
            "(unident\\. )*unident(ified|\\.)",
            "ZZZZunident.") %>%
  # Sort taxonomically. Put the "unident." at the end
  arrange(Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
  mutate_at(dplyr::vars(Kingdom:Species),
            str_replace,
            "ZZZZ",
            "") %>%
  mutate_at(dplyr::vars(Kingdom:Species),
            ~ factor(.) %>% fct_inorder)
```

```{r}
rand_palette <- partial(colorscheme$hex,
                        color_space = hcl_presets$colorblind_friendly,
                        force_mode = FALSE)
```

```{r}
summarize_taxonomy <- function(data, band_rank, group_rank, ...,
                               band_prop = 0.02, group_prop = 0.05, min_reads = 2,
                               print.table = if (is.null(get0("doctype"))) TRUE else
                                 if (doctype == "beamer") FALSE else
                                   TRUE) {
  band_rank <- ensym(band_rank)
  group_rank <- ensym(group_rank)
  classification <- tidyselect::vars_select(names(data), Kingdom:!!band_rank)
  data <- filter(data, Kingdom != "", n_reads >= min_reads, ...)
  data <- mutate_at(data, classification, fct_drop)

  # count the reads and OTUs for taxa at each rank

  if ("tech" %in% names(data)) data <- group_by(data, tech)
  for(rank in classification) {
    data <- group_by_at(data, rank, .add = TRUE)
    data <-
      mutate(data,
             !!(glue::glue("{rank}.otus")) := n(),
             !!(glue::glue("{rank}.reads")) := sum(n_reads))
  }
  class_counts <- glue::glue("{classification}.reads")
  data <- group_by_at(data, class_counts, .add = TRUE)
  data <- summarize(data,
                    `# OTUs` = n(),
                    `# reads` = sum(n_reads))
  while(is_grouped_df(data)) data <- ungroup(data)

  # Order each rank by number of reads, with most reads first.
  data <- arrange_at(data, class_counts, .funs = desc)
  data <- mutate_at(data, classification, fct_inorder)

  # Print the data
  if (print.table)
    print(knitr::kable(select(data, one_of(classification), `# OTUs`, `# reads`),
                       booktabs = TRUE))

  # Group taxa with a small fraction of the total OTUs and reads (on all techs) into "Other"
  if ("tech" %in% names(data)) data <- group_by(data, tech)

  data <- mutate(data,
                 total_OTUs = sum(`# OTUs`),
                 total_reads = sum(`# reads`))
  data <- group_by(data, !!group_rank, add = TRUE)
  data <- mutate(data,
                 group_OTUs = sum(`# OTUs`),
                 group_reads = sum(`# reads`))
  data <- group_by(data, !!band_rank, add = TRUE)
  data <- mutate(data,
                 band_OTUs = sum(`# OTUs`),
                 band_reads = sum(`# reads`))
  while(is_grouped_df(data)) data <- ungroup(data)

  data <- mutate_at(data, deparse(group_rank), as.character)
  data <- group_by(data, !!group_rank)
  data <- mutate(data,
                 group_other = !any(group_OTUs >= group_prop * total_OTUs,
                                     group_reads >= group_prop * total_reads))
  data <- ungroup(data)
  data <- mutate(data, !!group_rank := ifelse(group_other, "other", !!group_rank),
                 !!band_rank := as.character(!!band_rank))

  data <- group_by_at(data, classification)
  data <- mutate(data,
                 band_other = !any(band_OTUs >= band_prop * total_OTUs,
                                   band_reads >= band_prop * total_reads))
  data <- ungroup(data)
  data <- mutate(data, !!band_rank := ifelse(band_other, paste("other", !!group_rank), !!band_rank))
  data <- mutate_at(data, deparse(band_rank), str_replace, "other( other)+", "other")

  # data <- mutate_at(data, deparse(band_rank), fct_lump, w = data$`# OTUs`,
  #                   prop = band_prop, other_level = "other")
  # data <- mutate_at(data, deparse(group_rank), fct_lump, w = data$`# OTUs`,
  #                   prop = group_prop, other_level = "other")
  data <- mutate_at(data, classification, . %>% factor %>% fct_inorder)

  pal <- rand_palette(n = n_distinct(data[[deparse(band_rank)]]))
  data <- tidyr::gather(data, key = "count", value = "value", "# OTUs", "# reads")

  p1 <- ggplot(data, aes(x = !!group_rank, y = value, fill = !!band_rank)) +
    facet_wrap(facets = vars(count, tech), nrow = 2, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
    geom_col(position = "stack") +
    scale_x_discrete(deparse(group_rank), expand = c(0, 0)) +
    scale_fill_manual(values = pal) +
    theme(axis.text.x = element_text(angle = -20, vjust = 1, hjust = 0),
          legend.position = "bottom", legend.text = element_text(size = 8),
          legend.key.height = unit(5, "mm"))
}
```

# High level taxonomic ID
```{r, results='asis', fig.cap="OTU richness and read count for Phyla and Kingdoms, based on ITS2 metabarcoding"}
#debugonce(summarize_taxonomy)
summarize_taxonomy(all.otu, Phylum, Kingdom) + theme(legend.position = "right")
```

# Class-level Taxon ID
```{r, results = 'asis', fig.cap="OTU richness and read count for Classes and Phyla, based on ITS2 metabarcoding"}
summarize_taxonomy(all.otu, Class, Phylum) + theme(legend.position = "right")
```

# Order-level taxon ID

```{r, results = 'asis', fig.cap="OTU richness and read count for Orders and Phyla, based on ITS2 metabarcoding"}
summarize_taxonomy(all.otu, Order, Phylum) + theme(legend.position = "right")
```

# Genus-level taxon ID: Basidiomycota

```{r, results='asis', fig.cap="OTU richness and read count for Orders and Genera in Basidiomycota, based on ITS2 metabarcoding"}
summarize_taxonomy(all.otu, Genus, Order, Phylum == "Basidiomycota") + theme(legend.position = "right")
```

# Genus-level taxon ID: ECM

```{r, results='asis', fig.cap="OTU richness and read count for Orders and Genera, probable ECM species only, based on FUNGuild assignment of OTUs from ITS2 metabarcoding"}
summarize_taxonomy(all.otu, Genus, Order, str_detect(Guild, "Ectomycorrhizal")) + theme(legend.position = "right")
```

# Guild assignment
```{r}
all.otu %>%
  rename(`# reads` = n_reads) %>%
  mutate(`# OTUs` = 1) %>%
  mutate_at("Guild", str_replace, "^-$", "unknown") %>%
  mutate(Guild = fct_lump(Guild, prop = 0.02, w = `# reads`)) %>%
  tidyr::gather(key = "count", value = "value", "# OTUs", "# reads") %>%
  ggplot(aes(x = Guild, y = value, fill = Guild)) +
    facet_wrap(facets = vars(count, tech), nrow = 2, scales = "free_y") +
    geom_col(position = "stack") +
    scale_x_discrete("Guild", expand = c(0, 0)) +
    scale_fill_discrete( guide = "none") +
    theme(axis.text.x = element_text(angle = -20, vjust = 1, hjust = 0), legend.position = "bottom")
```

# Genus-level ID: Saprotrophs
```{r, results='asis', fig.cap="OTU richness and read count for Orders and Genera in Basidiomycota, based on short-read PacBio data"}
summarize_taxonomy(all.otu, Genus, Order, str_detect(Guild, "Undefined Saprotroph")) + theme(legend.position = "right")
```

# Sequence length distribution

```{r fig.width = 5, fig.height = 3}
 ggplot(all.otu, aes(SeqLen, ..density.., weight = n_reads, group = tech, color = tech)) +
  geom_freqpoly() +
  # facet_wrap(facets = vars(tech)) +
  scale_y_log10(name = "fraction of reads (log scale)") +
  scale_x_continuous(name = "Sequence length") +
  guides(color = guide_legend(direction = "horizontal", title = NULL)) +
  theme(legend.position = "bottom")
```

# Taxon ID: short PacBio sequences
```{r, results='asis', fig.cap="OTU richness and read count for Orders and Genera, representative sequence < 100bp"}
summarize_taxonomy(all.otu, Genus, Order, SeqLen < 100) + theme(legend.position = "right")
```

# Taxon ID: long PacBio sequences
\tiny
```{r, results='asis', fig.cap="OTU richness and read count for Orders and Genera, probable ECM species only, based on FUNGuild assignment of OTUs from ITS2 metabarcoding"}
all.otu %>% filter(SeqLen > 350) %>%
  select(Group, n_reads, SeqLen, Order:Species) %>%
  knitr::kable(booktabs = TRUE)
```