### 2019-04-26 writing/group_presentation.Rnw
Presentation for Anna's group

\documentclass[10]{beamer}
\include{pres_preamble}

\title{Turnover scale of soil fungal community composition}
\subtitle{Presentation for Anna's Group Meeting}
\author{Brendan Furneaux}
\date{26 April, 2019}

\begin{document}

\frame{\titlepage}

<<setup, include=FALSE>>=
knitr::opts_chunk$set(echo = FALSE, device = "tikz")
library(drake)
library(magrittr)
library(tidyverse)
library(ggplot2)
library(scales)
library(ggtree)
library(castor)
theme_set(theme_bw(base_size = 24) + theme(legend.direction = "vertical"))
loadd(bc_corr, bc_corr2, unif_corr, unif_corr2, qstats_join, datasets, longtree,
      taxon_ITS2_unite)
reads <- qstats_join %>%
  mutate_at("Region", factor,
            levels = c("ITS2", "ITS1", "short", "ITS", "LSU", "long")) %>%
  mutate_at("Dataset", factor,
            levels = c("short-ion", "short-pacbio", "long-pacbio")) %>%
  group_by(Dataset, Plate, Region)
reads_subset <- reads %>%
  do(if (nrow(.) > 10000) sample_n(., 10000) else .)
reverselog_trans <- function(base = exp(1)) {
    trans <- function(x) -log(x, base)
    inv <- function(x) base^(-x)
    trans_new(paste0("reverselog-", format(base)), trans, inv, 
              log_breaks(base = base), 
              domain = c(1e-100, Inf))
}
@

%\begin{frame}{Outline}
%\tableofcontents
%\end{frame}

\begin{frame}{Introduction}

\begin{itemize}
 \item Question: What is the scale of spatial (and temporal) autocorrelation in soil fungal community composition?
  \begin{itemize}
    \item Specifically in a tropical ECM forest.
  \end{itemize}
 \item Metabarcoding method comparison
\end{itemize}
\end{frame}

\begin{frame}{Upper Ouémé Forest Reserve}
\framesubtitle{Permanent study plots}
\vskip1.5em
\begincols
  \begincol{0.55\linewidth}
    \begin{figure}
      \begin{overprint}
      \onslide<1>
        \includegraphics{soudanian_benin}
      \onslide<2>
        \includegraphics{soudanian_benin_oueme}
      \onslide<3->
        \includegraphics{soudanian_benin_oueme_village}
      \end{overprint}
      \caption{Location of study sites in Bènin.}
    \end{figure}
  \endcol

  \begincol{0.45\linewidth}
  \begin{itemize}
    \tightlist
    \item<1-|alert@1> West Soudanian woodlands in central Bènin
    \item<1-|alert@1> 10° N Latitude
    \item<3-|alert@3> Two sites located near the villages Angaradebou and Gando
    \item<4-|alert@4> Dominated by \emph{Isoberlinia doka} (Caesalpinoideae; ECM host)
  \end{itemize}
  \endcol
\endcols
\end{frame}

\begin{frame}{Sampling}
  \begin{itemize}
    \item<1-|alert@1> \SI{25}{\m} transects.
    \item<1-|alert@1> Soil samples taken every 1 m.
    \item<2-|alert@2> $\SI{5}{\cm}\times\SI{5}{\cm}\times\SI{5}{\cm}$ sample, homogenize in plastic bag, then subsample \SI{50}{\mg}
    \item<3-|alert@3> Lysis \& preservation in field
    \item<4-|alert@4> Same locations in two years -- 2015 and 2016 (25 points × 2 years × 2 sites = 100 samples)
  \end{itemize}
\end{frame}

\begin{frame}{Sequencing}
  \framesubtitle{Which region?}
  \small
  \begincols
  \begincol{0.5\linewidth}
    \begin{block}{ITS2 (gITS7-ITS4)}
        \begin{itemize}
          \tightlist
          \item Internally transcribed spacer between 5.8S and 26/28S (LSU) subunits of ribosomal RNA
          \item Usually 250--350 bp.
          \item ITS --- standard DNA barcode for fungi.
          \item Relatively good species-level resolution, and good database support.
          \item Cannot be aligned across larger taxonomic groups.
          \item Groups with variant (or absent) 5.8S not sequenced.
        \end{itemize}
      \end{block}
    \endcol
    \begincol{0.5\linewidth}
      \begin{block}{Full ITS + partial LSU (ITS1-LR5)}
        \begin{itemize}
          \tightlist
          \item Full ITS plus the 5' end of LSU; 1500-2000 bp.
          \item ITS -- resolution to separate closely related taxa.
          \item LSU -- can be aligned across fungi.
          \item Primers are more universal than gITS7.
          \item Long amplicon with alternating variable and conserved regions --- susceptible to chimera formation and breakage.
        \end{itemize}
      \end{block}
    \endcol
  \endcols
\end{frame}

\begin{frame}{Sequencing}
  \framesubtitle{Which platform?}
  \begincols
  \begincol{0.5\linewidth}
    \begin{block}{IonTorrent}
      \begin{itemize}
        \tightlist
        \item Low cost per bp.
        \item Smaller minimum job than Illumina.
        \item Needs less DNA than PacBio.
        \item Bias against amplicons greater than 400 bp.
         \item Cannot sequence long ITS+LSU amplicon.
       \end{itemize}
     \end{block}
  \endcol
  \begincol{0.5\linewidth}
     \begin{block}{PacBio}
        \begin{itemize}
          \tightlist
          \item Higher cost per bp $\implies$ fewer samples or less sequencing depth for same budget.
          \item Requires more DNA.
          \item Good accuracy requires multiple circular reads.
        \end{itemize}
      \end{block}
  \endcol
  \endcols
  \centering

\end{frame}

\begin{frame}{Sequencing}
  \begin{block}{The plan}
  Three sequencing runs:
  \begin{itemize}
    \item Ion Proton S5, short amplicon
    \begin{itemize}
      \item Pool all samples on one chip
    \end{itemize}
    \item PacBio RSII, short amplicon
    \begin{itemize}
      \item Same amplicon library as IonTorrent.
      \item Two sequencing libraries, two SMRT cells per library.
    \end{itemize}
    \item PacBio RSII, long amplicon
    \begin{itemize}
      \item Two sequencing libraries, four SMRT cells per library.
    \end{itemize}
  \end{itemize}
  \end{block}
\end{frame}

\begin{frame}{Sequencing}
  \begin{block}{What happened}
  Three sequencing runs:
  \begin{itemize}
    \item \sout{IonTorrent, short amplicon}
    \begin{itemize}
      \item Accidentally used primers with IonTorrent adapters built in.
      \item Lost some of the multiplexing barcodes.
      \item Can't demultiplex the samples --- can only look at pool of all reads.
    \end{itemize}
    \item PacBio, short amplicon
    \begin{itemize}
      \item OK
    \end{itemize}
    \item PacBio, long amplicon
    \begin{itemize}
      \item OK
    \end{itemize}
  \end{itemize}
  \end{block}
\end{frame}

\begin{frame}{Bioinformatics}
  \begin{itemize}
    \item Redo RSII circular consensus basecalling using new algorithm (Arrow).
    \item Demultiplex and trim primers.
    \item Cut out regions (ITS1, ITS2, full ITS, LSU) using ITSx.
    \item Quality filter.
  \end{itemize}
\end{frame}

\begin{frame}{Expected error rate}
\framesubtitle{After \texttt{Arrow}}

<<errorrate, fig.width = 12>>=
ggplot(reads_subset, aes(x = erate,
                  linetype = Dataset,
                  group = paste(Dataset, Region),
                  color = Region)) +
  geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
  scale_x_continuous(name = "Expected number of errors per base",
                     trans = reverselog_trans(10),
                     limits = c(NA, 1e-5),
                     oob = squish) +
  scale_y_continuous(name = "Fraction passing") +
  scale_color_viridis_d(option = "C") +
  scale_linetype_discrete()
@
\end{frame}

\begin{frame}{Bioinformatics}
\framesubtitle{continued...}
  \begin{itemize}
    \item Find amplicon sequence variants (ASVs) with \texttt{dada2}.
      \begin{itemize}
        \item Doesn't work well on full long amplicons -- algorithm depends on having a large number of error-free reads.
        \item Works on the smaller regions.
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Expected total number of errors}
\framesubtitle{After \texttt{Arrow}}

<<totalerror, fig.width = 12>>=
ggplot(reads_subset, aes(x = eexp,
                  linetype = Dataset,
                  group = paste(Dataset, Region),
                  color = Region)) +
  geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
  scale_x_continuous(name = "Expected number of errors",
                     trans = reverselog_trans(10),
                     limits = c(NA, 0.1),
                     oob = squish) +
  scale_y_continuous(name = "Fraction passing") +
  scale_color_viridis_d(option = "C") +
  scale_linetype_discrete()
@
\end{frame}

\begin{frame}{Bioinformatics}
\framesubtitle{continued...}
  \begin{itemize}
    \item Find amplicon sequence variants (ASVs) with \texttt{dada2}.
      \begin{itemize}
        \item Doesn't work well on full long amplicons -- algorithm depends on having a large number of error-free reads.
        \item Works on the smaller regions.
      \end{itemize}
    \item For each ITS2 ASV, align the whole long read and find the consensus.
      \begin{itemize}
        \item There were a lot of chimeras (15--25\%).
        \item Remove "outliers" from the alignments using \texttt{odseq}.
        \item Remaining chimeras are too few to change the consensus.
      \end{itemize}
    \item Align with \texttt{DECIPHER}
      \begin{itemize}
        \item Progressive aligner with inferred RNA secondary structure.
      \end{itemize}
    \item \texttt{RAxML} tree
  \end{itemize}
\end{frame}

\begin{frame}{Taxonomy assignment}
Both \texttt{dada2} and \texttt{DECIPHER} can do taxonomy assignment.
\begincols
\begincol{0.5\framewidth}
\begin{block}{dada2}
  \begin{itemize}
    \item Naïve Bayesian classifier.
    \item Uses a lot of memory.
    \item Needs every rank to be filled (\emph{incertae sedis}, "unidentified").
    \item Seems to put thins in places that don't match the tree.
  \end{itemize}
\end{block}
\endcol

\begincol{0.5\framewidth}
\begin{block}{DECIPHER}
  \begin{itemize}
    \item Some kind of machine learning algorithm.
    \item Uses less memory.
    \item Allows uneven classification depths.
    \item Very low bootstraps, especially for ITS.
  \end{itemize}
\end{block}
\endcol
\endcols

UNITE for ITS, RDP for LSU...  But there is no large database that has both for the same specimens!

I'm still working on it...
\end{frame}

\begin{frame}{Tree}
\framesubtitle{Kingdoms}
<<kingtree, fig.width = 12, message=FALSE>>=
seqhash <- function(seq, len = 8) UseMethod("seqhash")
seqhash.character <- function(seq, len = 8) {
  h <- purrr::map_chr(seq, digest::digest)
  stringr::str_sub(h, end = len)
}
seqhash.XStringSet <- function(seq, len = 8) {
  seqhash.character(as.character(seq))
}
treetax <- taxon_ITS2_unite %>%
  mutate_at("seq", seqhash) %>%
  select(-Taxonomy, -nreads) 
outgroups <- treetax %>% filter(!is.na(Kingdom), Kingdom %in% c("Alveolata", "Chromista", "Plantae"), seq %in% longtree$tip.label)
p <- ggtree(tr = longtree)
longtree <- reroot(longtree, MRCA(p, outgroups$seq))

ggtree(groupOTU(longtree, split(treetax$seq, treetax$Kingdom))) +
  aes(color = group) + scale_color_brewer(type = "qual", palette = "Paired") +
  theme(legend.position = NULL)
@
\end{frame}

\begin{frame}{Tree}
\framesubtitle{Phyla}
<<phytree, fig.width = 12, message=FALSE>>=
ggtree(groupOTU(longtree, split(treetax$seq, treetax$Phylum))) +
  aes(color = group) + scale_color_brewer(type = "qual", palette = "Paired") +
  theme(legend.position = NULL)
@
\end{frame}

\begin{frame}{Tree}
\framesubtitle{Families --- Basidiomycota}
<<basidtree, fig.width = 12, message=FALSE>>=
basids <- treetax %>% filter(!is.na(Class),
                             Class %in% c("Geminibasidiomycetes",
                                         "Tremellomycetes",
                                         "Exobasidiomycetes",
                                         "Microbotryomycetes"),
                             seq %in% longtree$tip.label)

basidtree <- phylobase::phylo4(longtree) %>%
  phylobase::subset(mrca = basids$seq) %>%
  as("phylo")

basids2 <- treetax %>% filter(seq %in% basidtree$tip.label)

ggtree(groupOTU(basidtree, split(basids2$seq, basids2$Order)), layout = "circular") +
  aes(color = group) +
  scale_color_brewer(type = "qual", palette = "Paired", limits = unique(na.omit(basids2$Order))) +
  theme(legend.position = NULL)
@
\end{frame}

\begin{frame}{Correlogram}
<<correlogram, fig.width = 12>>=
bind_rows(as.data.frame(bc_corr$mantel.res) %>%
            mutate(tdist = "same", metric = "bray"),
          as.data.frame(bc_corr2$mantel.res) %>%
            mutate(tdist = "diff", metric = "bray"),
          as.data.frame(unif_corr$mantel.res) %>%
            mutate(tdist = "same", metric = "unifrac"),
          as.data.frame(unif_corr2$mantel.res) %>%
            mutate(tdist = "diff", metric = "unifrac")) %>%
  mutate(Significant = `Pr(corrected)` < 0.05) %>%
  ggplot(aes(class.index, Mantel.cor, group = tdist, color = tdist, shape = Significant)) +
  geom_point(size = 3) +
  geom_line(alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
  # geom_line(aes(y = exp(predict(bc_fit)))) +
  scale_y_continuous(limits = c(-0.05, NA), name = "Spatial Distance correlation") +
  scale_color_discrete(name = "Year") +
  xlab("distance (m)") +
  facet_wrap("metric") +
  scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1))
@
\end{frame}

\end{document}