### 2019-06-05 writing/coverage.Rmd

---
title: "Coverage tests"
author: "Brendan Furneaux"
date: "2/22/2019"
output: 
  beamer_presentation: 
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      external = FALSE,
                      device = "tikz")
library(magrittr)
library(drake)
library(ggplot2)
library(tidyverse)
theme_set(theme_bw(base_size = 24))

bigmaps <- drake::cached() %>%
  str_subset("dada_map_pb_500") %>%
  map(readd, character_only = TRUE)

pm <- readd(platemap) %>%
  mutate_at("Primer.Pair", tolower) %>%
  mutate_at(c("Year", "Site", "Qual", "Plate", "Well", "Primer.Pair"), factor)

st <- readd(big_seq_table_ITS2) %>%
  set_colnames(map_chr(colnames(.), digest::digest) %>% substr(1,8)) %>%
  as_tibble(rownames = "file") %>%
  tidyr::extract(
    col = "file",
    into = c("Seq.Run", "Plate", "Well", "Direction", "Region"),
    regex = "([:alpha:]+_\\d+)_(\\d+)-([A-H]1?\\d)([fr]?)-([:alnum:]+)\\.qfilt\\.fastq\\.gz") %>%
  group_by(Seq.Run, Plate, Well, Region) %>%
  summarize_if(is.numeric, sum) %>%
  ungroup() %>%
  left_join(readd(datasets)) %>%
  mutate(Primer.Pair = paste(str_replace(Forward, "_tag.*$", ""),
                             str_replace(Reverse, "_tag.*$", ""),
                             sep = "_")) %>%
  select(Seq.Run, Plate, Well, Region, Primer.Pair, everything()) %>%
  mutate_at(c("Seq.Run", "Region", "Primer.Pair"), factor) %>%
  mutate_at("Well", factor, levels = levels(pm$Well)) %>%
  mutate_at("Plate", factor, levels = levels(pm$Plate)) %>%
  gather(key = "ASV", value = "Reads",
         -one_of(c("Seq.Run", "Plate", "Well", "Region", "Primer.Pair", "Dataset", "Tech", "Runs", "Forward", "Reverse", "Regions", "PlateKey", "Homopolymer.Gap.Penalty", "Band.Size", "Pool"))) %>%
  mutate_at("Reads", as.integer) %>%
  filter(Reads > 0L)
```

# Accumulation curves

```{r coveragesample, results='asis'}
right_join(pm, st) %>%
    filter(!is.na(Year), Qual == "") %>%
    group_by(Site, X, Year, Seq.Run, ASV) %>%
    summarize(Reads = 1) %>%
    group_by(Site, Year, Seq.Run, ASV) %>%
    summarize(Reads = sum(Reads)) %>%
    spread(key = ASV, value = Reads, fill = 0L) %>%
    group_by(Site, Year) %>%
    group_walk(~ {cat("\n\n## Coverage by # samples | ",
                      as.character(.y$Site),
                      as.character(.y$Year), "\n\n")
      column_to_rownames(.x, "Seq.Run") %>%
                 as.matrix() %>%
                 t %>%
                 rbind(25, .) %>%
                 iNEXT::iNEXT(datatype = "incidence_freq") %>%
                 {iNEXT::ggiNEXT(.) +
                     xlab("Number of samples") +
                     ylab("ASV diversity") +
                     scale_fill_discrete(limits = levels(.x[["Seq.Run"]]),
                                         drop = FALSE) +
                     scale_color_discrete(limits = levels(.x[["Seq.Run"]]),
                                         drop = FALSE) +
                     scale_shape_discrete(limits = levels(.x[["Seq.Run"]]),
                                         drop = FALSE)} %>%
                 plot})
```

```{r coveragereads, results='asis'}
right_join(pm, st) %>%
    filter(!is.na(Year), Qual == "") %>%
    group_by(Site, Year, Seq.Run, ASV) %>%
    summarize(Reads = sum(Reads)) %>%
    spread(key = ASV, value = Reads, fill = 0L) %>%
    group_by(Site, Year) %>%
    group_walk(~ {cat("\n\n## Coverage by # reads |",
                      as.character(.y$Site),
                      as.character(.y$Year), "\n\n")
      column_to_rownames(.x, "Seq.Run") %>%
                 as.matrix() %>%
                 t %>%
                 iNEXT::iNEXT(datatype = "abundance") %>%
                 {iNEXT::ggiNEXT(.) +
                     xlab("Number of reads") +
                     ylab("ASV diversity") +
                     scale_fill_discrete(limits = levels(.x[["Seq.Run"]]),
                                         drop = FALSE) +
                     scale_color_discrete(limits = levels(.x[["Seq.Run"]]),
                                         drop = FALSE) +
                     scale_shape_discrete(limits = levels(.x[["Seq.Run"]]),
                                         drop = FALSE)} %>%
                 plot})
```
```{r coveragesamplereads, results='asis'}
d <- right_join(pm, st) %>%
    filter(!is.na(Year)) %>%
    group_by(Site, Year, X, Qual, Seq.Run, ASV) %>%
    summarize(Reads = sum(Reads)) %>%
    spread(key = ASV, value = Reads, fill = 0L) %>%
    group_by(Site, X)
cats <- ungroup(d) %>%
  select(Year, Seq.Run, Qual) %>%
  unique() %>%
  glue::glue_data("{Year}_{Seq.Run}_{Qual}")

group_walk(d, 
           ~ {cat("\n\n## Coverage by # reads |",
                  as.character(.y$Site),
                  "sample",
                  as.character(.y$X),
                  "\n\n")
             unite(.x, "Seq.Run", Year, Seq.Run, Qual) %>%
               column_to_rownames("Seq.Run") %>%
               as.matrix() %>%
               t %>%
               iNEXT::iNEXT(datatype = "abundance") %>%
               {iNEXT::ggiNEXT(.) +
                   xlab("Number of reads") +
                   ylab("ASV diversity") +
                   scale_fill_discrete(limits = cats,
                                       drop = FALSE) +
                   scale_color_discrete(limits = cats,
                                        drop = FALSE) +
                   scale_shape_discrete(limits = cats,
                                        drop = FALSE)} %>%
               plot})
```

# Resampling

## Read counts per site×sample×year

```{r}
right_join(pm, st) %>%
  filter(startsWith(as.character(Seq.Run), "pb"),
         !is.na(Year)) %>%
  group_by(Year, Site, X, Qual, Dataset) %>%
  summarize(Ntot = sum(Reads)) %>%
  ggplot(aes(x = Ntot, group = Dataset, color = Dataset)) +
  geom_line(mapping = aes(y = 1 - ..y..), stat = "ecdf") +
  xlab("Number of reads") +
  ylab("Fraction of samples")
```

## Read counts per site×sample

```{r}
right_join(pm, st) %>%
  filter(startsWith(as.character(Seq.Run), "pb"),
         !is.na(Year)) %>%
  group_by(Site, X, Qual, Dataset) %>%
  summarize(Ntot = sum(Reads)) %>%
  ggplot(aes(x = Ntot, group = Dataset, color = Dataset)) +
  geom_line(mapping = aes(y = 1 - ..y..), stat = "ecdf") +
  xlab("Number of reads") +
  ylab("Fraction of samples")
```

## New vs. expired buffer

 - in 2016, there was not enough unexpired field lysis buffer for all samples
 - sampled with expired buffer at all locations
 - sampled with unexpired buffer at approx. 1/3 of all locations.

## Reads: new vs. expired buffer

```{r}
right_join(pm, st) %>%
  filter(!is.na(Year), Qual != "*") %>%
  group_by(Dataset, Year, X, Site) %>%
  filter("" %in% Qual & "X" %in% Qual) %>%
  group_by(Dataset, Year, X, Site, Qual) %>%
  summarize(Reads = sum(Reads)) %>%
  mutate_at("Qual", forcats::fct_recode, Expired = "", New = "X") %>%
  spread(key = Qual, value = Reads, fill = 0L) %>%
  ggplot(aes(x = Expired, y = New, color = Dataset)) +
  geom_point() +
  geom_abline(slope = 1, linetype = 2, alpha = 0.5) +
  coord_fixed() +
  xlab("Expired buffer") +
  ylab("New buffer")
```

## Reads: new vs expired buffer

```{r}
right_join(pm, st) %>%
  filter(!is.na(Year), Qual != "*") %>%
  group_by(Dataset, Year, X, Site) %>%
  filter("" %in% Qual & "X" %in% Qual) %>%
  group_by(Dataset, Year, X, Site, Qual) %>%
  summarize(Reads = sum(Reads)) %>%
  mutate_at("Qual", forcats::fct_recode, Expired = "", New = "X") %>%
  spread(key = Qual, value = Reads, fill = 0L) %>%
  filter(New > 20, Expired > 20) %>%
  mutate(Ratio = New / Expired) %>%
  ggplot(aes(y = Ratio, x = Dataset)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, alpha = 0.5, linetype = 2)
```