### 2019-06-10 writing/tech_compare.Rmd
Add ASV comparison, fractions by read, secondary axis with community fraction

---
title: "Read quality check"
output:
  beamer_presentation:
    latex_engine: xelatex
---

```{r setup, include = FALSE}
library(magrittr)
library(tidyverse)
library(here)
library(glue)
library(Biostrings)
library(ShortRead)
library(ggplot2)
library("scales")
library(drake)
knitr::opts_chunk$set(echo = FALSE, device = "tikz")
theme_set(theme_bw(base_size = 24) + theme(legend.direction = "vertical"))
```

# Read number from PacBio vs. IonTorrent ASVs (ITS2)
```{r}
loadd(big_seq_table_ITS2)
asv_table <- big_seq_table_ITS2 %>%
  tibble::as_tibble(rownames = "filename") %>%
  tidyr::gather(key = "seq", value = "reads", -1) %>%
  dplyr::filter(reads >= 1) %>%
  tidyr::extract(col = "filename",
                 into = c("seq.run", "plate", "well", "dir", "region"),
                 regex = "([a-z]+_\\d+)_(\\d+)-([A-H]1?[0-9])([fr]?)-([:alnum:]+).+") %>%
  dplyr::filter(seq.run %in% c("is_057", "pb_500")) %>%
  dplyr::group_by(seq.run, seq) %>%
  dplyr::summarize(reads = sum(reads)) %>%
  tidyr::spread(key = seq.run, value = reads, fill = 0)

asv_table %>% 
  mutate(is_057 = ifelse(is_057 == 0, 0.6, is_057),
                    pb_500 = ifelse(pb_500 == 0, 0.6, pb_500)) %>%
  ggplot(aes(x = is_057, y = pb_500)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  xlab("Ion Torrent Reads") +
  ylab("PacBio Reads")

```

# Read number from PacBio vs. IonTorrent ASV-OTUs (ITS2)

```{r}
otu <- read_tsv(here("data/clusters/ITS2.table"),
                col_types = "ciii") %>% select(is_057, pb_500) %>%
  filter(rowSums(.) > 0)
otu %>%
  mutate(is_057 = ifelse(is_057 == 0, 0.6, is_057),
         pb_500 = ifelse(pb_500 == 0, 0.6, pb_500)) %>%
  ggplot(aes(x = is_057, y = pb_500)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  xlab("Ion Torrent Reads") +
  ylab("PacBio Reads")
```

# IonTorrent ITS2 ASVs: coverage by PacBio long reads
```{r}
asv_table %>%
  arrange(desc(is_057)) %>%
  filter(is_057 > 0) %>%
  ggplot(aes(y = is_057, x = seq_along(is_057), fill = pb_500 > 0)) +
  geom_col(width = 1) +
  xlab("ASVs") +
  ylab("reads") +
  scale_fill_discrete(name = "PacBio") +
  scale_y_continuous(sec.axis = sec_axis(~ ./sum(.), name = "frac. of comm."))
```
`r asv_table %>% filter(is_057 > 0) %>% {formatC(sum(.$pb_500 > 0) / length(.$is_057)*100, digits = 1, flag = "0", format = "f")}`% of IonTorrent ASVs (`r asv_table %>% {formatC(sum(.$is_057[.$pb_500 > 0]) / sum(.$is_057) * 100, digits = 1, flag = "0", format = "f")}`% of reads) also include at least one PacBio read.

# IonTorrent ITS2 ASV-OTUs: coverage by PacBio long reads
```{r}
otu %>%
  arrange(desc(is_057)) %>%
  filter(is_057 > 0) %>%
  ggplot(aes(y = is_057, x = seq_along(is_057), fill = pb_500 > 0)) +
  geom_col(width = 1) +
  xlab("OTUs") +
  ylab("reads") +
  scale_fill_discrete(name = "PacBio") +
  scale_y_continuous(sec.axis = sec_axis(~ ./sum(.), name = "frac. of comm."))
```
`r otu %>% filter(is_057 > 0) %>% {formatC(sum(.$pb_500 > 0) / length(.$is_057)*100, digits = 1, flag = "0", format = "f")}`% of IonTorrent OTUs (`r otu %>% {formatC(sum(.$is_057[.$pb_500 > 0]) / sum(.$is_057) * 100, digits = 1, flag = "0", format = "f")}`% of reads) also include at least one PacBio ASV.

# PacBio ITS2 ASVs: coverage by Ion Torrent
```{r}
asv_table %>%
  arrange(desc(pb_500)) %>%
  filter(pb_500 > 0) %>%
  ggplot(aes(y = pb_500, x = seq_along(pb_500), fill = is_057 > 0)) +
  geom_col(width = 1) +
  xlab("ASVs") +
  ylab("reads") +
  scale_fill_discrete(name = "Ion Torrent") +
  scale_y_continuous(sec.axis = sec_axis(~ ./sum(.), name = "frac. of comm."))
```
`r asv_table %>% filter(pb_500 > 0) %>% {formatC(sum(.$is_057 > 0) / length(.$pb_500)*100, digits = 1, flag = "0", format = "f")}`% of PacBio ASVs (`r asv_table %>% {formatC(sum(.$pb_500[.$is_057 > 0]) / sum(.$pb_500) * 100, digits = 1, flag = "0", format = "f")}`% of reads) also include at least one IonTorrent read

# PacBio ITS2 ASV-OTUs: coverage by Ion Torrent
```{r}
otu %>%
  arrange(desc(pb_500)) %>%
  filter(pb_500 > 0) %>%
  ggplot(aes(y = pb_500, x = seq_along(pb_500), fill = is_057 > 0)) +
  geom_col(width = 1) +
  xlab("OTUs") +
  ylab("reads") +
  scale_fill_discrete(name = "Ion Torrent") +
  scale_y_continuous(sec.axis = sec_axis(~ ./sum(.), name = "frac. of comm."))
```
`r otu %>% filter(pb_500 > 0) %>% {formatC(sum(.$is_057 > 0) / length(.$pb_500)*100, digits = 1, flag = "0", format = "f")}`% of PacBio OTUs (`r otu %>% {formatC(sum(.$pb_500[.$is_057 > 0]) / sum(.$pb_500) * 100, digits = 1, flag = "0", format = "f")}`% of reads) also include at least one IonTorrent ASV