### 2019-08-12 writing/MSA2019_presentation.Rnw
add logo, comment out unneeded slides

\documentclass[10]{beamer}
\include{pres_preamble}

\title{Comparison of long and short amplicon metabarcoding for determination of fungal species diversity and spatial turnover in an ectomycorrhizal West African woodland}
\author{\mbox{\textbf{Brendan Furneaux}\inst{1}} \and \mbox{Mohammad Bahram\inst{2}} \and \mbox{Nourou Yorou\inst{3}} \and \mbox{Anna Rosling\inst{1}} \and \mbox{Martin Ryberg\inst{1}}}
\institute{\inst{1}Uppsala University \and \inst{2}Swedish University of Agricultural Sciences \and \inst{3}University of Parakou}
\date{13 August, 2019}
\logo{\includegraphics[width=2.5cm]{Uppsala_University_logo}}

\begin{document}

\frame{\titlepage}

<<setup, include=FALSE>>=
knitr::opts_chunk$set(echo = FALSE, device = "tikz")
library(drake)
library(magrittr)
library(tidyverse)
library(ggplot2)
library(scales)
library(ggtree)
library(castor)
theme_set(theme_bw(base_size = 24) + theme(legend.direction = "vertical"))
loadd(bc_corr, bc_corr2, unif_corr, unif_corr2, qstats_join, datasets, longtree,
      taxon_LSUcons_rdp)
reads <- qstats_join %>%
  mutate_at("Region", factor,
            levels = c("ITS2", "ITS1", "short", "ITS", "LSU", "long")) %>%
  mutate_at("Dataset", factor,
            levels = c("short-ion", "short-pacbio", "long-pacbio")) %>%
  group_by(Dataset, Plate, Region)
reads_subset <- reads %>%
  do(if (nrow(.) > 10000) sample_n(., 10000) else .)
reverselog_trans <- function(base = exp(1)) {
    trans <- function(x) -log(x, base)
    inv <- function(x) base^(-x)
    trans_new(paste0("reverselog-", format(base)), trans, inv, 
              log_breaks(base = base), 
              domain = c(1e-100, Inf))
}
@

% # Metabarcoding
% 
%   - Cloning + Sanger -> 454 -> Illumina
%   - For fungi, usually ITS2
%       - short
%       - intraspecific variation in most groups
%       - good database support (Unite 8.0 - 817 130 fungal sequences)
%       - ITS is "official barcode for fungi" \cite{schoch2012}
%       - fungal specific primers
%   - problems
%       - not alignable
%       - cannot place unknown sequences
%       - variant 5.8S
%       - variable length
%   
%   - LSU (26S/28S)
%       - alternating highly conserved domains and variable regions
%       - widely used for phylogenies
%       - universal primers
%       - less variable length
%       - more conserved than ITS
%       - not as good database support (RDP 11 fungal LSU - 125 525, Silva 132 LSU 907 382 of which )
%   - SSU (18S)
%       - alternating highly conserved and variable regions
%       - more conserved than LSU
%       - used for barcoding prokaryotes and "protists"
%       - good database support (Silva SSU), but not as good for fungi

%\begin{frame}{Outline}
%\tableofcontents
%\end{frame}

\begin{frame}{Introduction}

\begin{itemize}
 \item Question: What is the scale of spatial (and temporal) autocorrelation in soil fungal community composition?
  \begin{itemize}
    \item Specifically in a tropical ECM forest.
  \end{itemize}
 \item Metabarcoding method comparison
\end{itemize}
\end{frame}

\begin{frame}{Upper Ouémé Forest Reserve}
\framesubtitle{Permanent study plots}
\vskip1.5em
\begincols
  \begincol{0.55\linewidth}
    \begin{figure}
      \begin{overprint}
      \onslide<1>
        \includegraphics{soudanian_benin}
      \onslide<2>
        \includegraphics{soudanian_benin_oueme}
      \onslide<3->
        \includegraphics{soudanian_benin_oueme_village}
      \end{overprint}
      \caption{Location of study sites in Bènin.}
    \end{figure}
  \endcol

  \begincol{0.45\linewidth}
  \begin{itemize}
    \tightlist
    \item<1-|alert@1> West Soudanian woodlands in central Bènin
    \item<1-|alert@1> 10° N Latitude
    \item<3-|alert@3> Two sites located near the villages Angaradebou and Gando
    \item<4-|alert@4> Dominated by \emph{Isoberlinia doka} (Caesalpinoideae; ECM host)
  \end{itemize}
  \endcol
\endcols
\end{frame}

\begin{frame}{Sampling}
  \begin{itemize}
    \item<1-|alert@1> \SI{25}{\m} transects.
    \item<1-|alert@1> Soil samples taken every 1 m.
    \item<2-|alert@2> $\SI{5}{\cm}\times\SI{5}{\cm}\times\SI{5}{\cm}$ sample, homogenize in plastic bag, then subsample \SI{50}{\mg}
    \item<3-|alert@3> Lysis \& preservation in field
    \item<4-|alert@4> Same locations in two years -- 2015 and 2016 (25 points × 2 years × 2 sites = 100 samples)
  \end{itemize}
\end{frame}

\begin{frame}{Sequencing}
  \framesubtitle{Which region?}
  \small
  \begincols
  \begincol{0.5\linewidth}
    \begin{block}{ITS2 (gITS7-ITS4)}
        \begin{itemize}
          \tightlist
          \item Internally transcribed spacer between 5.8S and 26/28S (LSU) subunits of ribosomal RNA
          \item Usually 250--350 bp.
          \item ITS --- standard DNA barcode for fungi.
          \item Relatively good species-level resolution, and good database support.
          \item Cannot be aligned across larger taxonomic groups.
          \item Groups with variant (or absent) 5.8S not sequenced.
        \end{itemize}
      \end{block}
    \endcol
    \begincol{0.5\linewidth}
      \begin{block}{Full ITS + partial LSU (ITS1-LR5)}
        \begin{itemize}
          \tightlist
          \item Full ITS plus the 5' end of LSU; 1500-2000 bp.
          \item ITS -- resolution to separate closely related taxa.
          \item LSU -- can be aligned across fungi.
          \item Primers are more universal than gITS7.
          \item Long amplicon with alternating variable and conserved regions --- susceptible to chimera formation and breakage.
        \end{itemize}
      \end{block}
    \endcol
  \endcols
\end{frame}

% \begin{frame}{Sequencing}
%   \framesubtitle{Which platform?}
%   \begincols
%   \begincol{0.5\linewidth}
%     \begin{block}{IonTorrent}
%       \begin{itemize}
%         \tightlist
%         \item Low cost per bp.
%         \item Smaller minimum job than Illumina.
%         \item Needs less DNA than PacBio.
%         \item Bias against amplicons greater than 400 bp.
%          \item Cannot sequence long ITS+LSU amplicon.
%        \end{itemize}
%      \end{block}
%   \endcol
%   \begincol{0.5\linewidth}
%      \begin{block}{PacBio}
%         \begin{itemize}
%           \tightlist
%           \item Higher cost per bp $\implies$ fewer samples or less sequencing depth for same budget.
%           \item Requires more DNA.
%           \item Good accuracy requires multiple circular reads.
%         \end{itemize}
%       \end{block}
%   \endcol
%   \endcols
%   \centering
% 
% \end{frame}
% 
% \begin{frame}{Sequencing}
%   \begin{block}{The plan}
%   Three sequencing runs:
%   \begin{itemize}
%     \item Ion Proton S5, short amplicon
%     \begin{itemize}
%       \item Pool all samples on one chip
%     \end{itemize}
%     \item PacBio RSII, short amplicon
%     \begin{itemize}
%       \item Same amplicon library as IonTorrent.
%       \item Two sequencing libraries, two SMRT cells per library.
%     \end{itemize}
%     \item PacBio RSII, long amplicon
%     \begin{itemize}
%       \item Two sequencing libraries, four SMRT cells per library.
%     \end{itemize}
%   \end{itemize}
%   \end{block}
% \end{frame}
% 
% \begin{frame}{Sequencing}
%   \begin{block}{What happened}
%   Three sequencing runs:
%   \begin{itemize}
%     \item \sout{IonTorrent, short amplicon}
%     \begin{itemize}
%       \item Accidentally used primers with IonTorrent adapters built in.
%       \item Lost some of the multiplexing barcodes.
%       \item Can't demultiplex the samples --- can only look at pool of all reads.
%     \end{itemize}
%     \item PacBio, short amplicon
%     \begin{itemize}
%       \item OK
%     \end{itemize}
%     \item PacBio, long amplicon
%     \begin{itemize}
%       \item OK
%     \end{itemize}
%   \end{itemize}
%   \end{block}
% \end{frame}

\begin{frame}{Bioinformatics}
  \begin{itemize}
    \item Redo RSII circular consensus basecalling using new algorithm (Arrow).
    \item Demultiplex and trim primers.
    \item Cut out regions (ITS1, ITS2, full ITS, LSU) using ITSx.
    \item Quality filter.
  \end{itemize}
\end{frame}

\begin{frame}{Expected error rate}
\framesubtitle{After \texttt{Arrow}}

% <<errorrate, fig.width = 12>>=
% ggplot(reads_subset, aes(x = erate,
%                   linetype = Dataset,
%                   group = paste(Dataset, Region),
%                   color = Region)) +
%   geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
%   scale_x_continuous(name = "Expected number of errors per base",
%                      trans = reverselog_trans(10),
%                      limits = c(NA, 1e-5),
%                      oob = squish) +
%   scale_y_continuous(name = "Fraction passing") +
%   scale_color_viridis_d(option = "C") +
%   scale_linetype_discrete()
% @
\end{frame}

\begin{frame}{Bioinformatics}
\framesubtitle{continued...}
  \begin{itemize}
    \item Find amplicon sequence variants (ASVs) with \texttt{dada2}.
      \begin{itemize}
        \item Doesn't work well on full long amplicons -- algorithm depends on having a large number of error-free reads.
        \item Works on the smaller regions.
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Expected total number of errors}
\framesubtitle{After \texttt{Arrow}}

% <<totalerror, fig.width = 12>>=
% ggplot(reads_subset, aes(x = eexp,
%                   linetype = Dataset,
%                   group = paste(Dataset, Region),
%                   color = Region)) +
%   geom_line(aes(y = 1 - ..y..), stat = "ecdf") +
%   scale_x_continuous(name = "Expected number of errors",
%                      trans = reverselog_trans(10),
%                      limits = c(NA, 0.1),
%                      oob = squish) +
%   scale_y_continuous(name = "Fraction passing") +
%   scale_color_viridis_d(option = "C") +
%   scale_linetype_discrete()
% @
\end{frame}

\begin{frame}{Bioinformatics}
\framesubtitle{continued...}
  \begin{itemize}
    \item Find amplicon sequence variants (ASVs) with \texttt{dada2}.
      \begin{itemize}
        \item Doesn't work well on full long amplicons -- algorithm depends on having a large number of error-free reads.
        \item Works on the smaller regions.
      \end{itemize}
    \item For each ITS2 ASV, align the whole long read and find the consensus.
      \begin{itemize}
        \item There were a lot of chimeras (15--25\%).
        \item Remove "outliers" from the alignments using \texttt{odseq}.
        \item Remaining chimeras are too few to change the consensus.
      \end{itemize}
    \item Align with \texttt{DECIPHER}
      \begin{itemize}
        \item Progressive aligner with inferred RNA secondary structure.
      \end{itemize}
    \item \texttt{RAxML} tree
  \end{itemize}
\end{frame}

\begin{frame}{Taxonomy assignment}
Both \texttt{dada2} and \texttt{DECIPHER} can do taxonomy assignment.
\begincols
\begincol{0.5\framewidth}
\begin{block}{dada2}
  \begin{itemize}
    \item Naïve Bayesian classifier.
    \item Uses a lot of memory.
    \item Needs every rank to be filled (\emph{incertae sedis}, "unidentified").
    \item Seems to put thins in places that don't match the tree.
  \end{itemize}
\end{block}
\endcol

\begincol{0.5\framewidth}
\begin{block}{DECIPHER}
  \begin{itemize}
    \item Some kind of machine learning algorithm.
    \item Uses less memory.
    \item Allows uneven classification depths.
    \item Very low bootstraps, especially for ITS.
  \end{itemize}
\end{block}
\endcol
\endcols

UNITE for ITS, RDP for LSU...  But there is no large database that has both for the same specimens!

I'm still working on it...
\end{frame}

\begin{frame}{Tree}
\framesubtitle{Kingdoms}
<<kingtree, fig.width = 12, message=FALSE>>=
seqhash <- function(seq, len = 8) UseMethod("seqhash")
seqhash.character <- function(seq, len = 8) {
  h <- purrr::map_chr(seq, digest::digest)
  stringr::str_sub(h, end = len)
}
seqhash.XStringSet <- function(seq, len = 8) {
  seqhash.character(as.character(seq))
}
treetax <- taxon_LSUcons_rdp %>%
  mutate_at("seq", seqhash) %>%
  select(-Taxonomy, -nreads) 
outgroups <- treetax %>% filter(!is.na(Kingdom), Kingdom %in% c("Alveolata", "Chromista", "Plantae"), seq %in% longtree$tip.label)
p <- ggtree(tr = longtree)
longtree <- reroot(longtree, MRCA(p, outgroups$seq))

ggtree(groupOTU(longtree, split(treetax$seq, treetax$Kingdom))) +
  aes(color = group) + scale_color_brewer(type = "qual", palette = "Paired") +
  theme(legend.position = NULL)
@
\end{frame}

\begin{frame}{Tree}
\framesubtitle{Phyla}
<<phytree, fig.width = 12, message=FALSE>>=
ggtree(groupOTU(longtree, split(treetax$seq, treetax$Phylum))) +
  aes(color = group) + scale_color_brewer(type = "qual", palette = "Paired") +
  theme(legend.position = NULL)
@
\end{frame}

\begin{frame}{Tree}
\framesubtitle{Families --- Basidiomycota}
<<basidtree, fig.width = 12, message=FALSE>>=
basids <- treetax %>% filter(!is.na(Class),
                             Class %in% c("Geminibasidiomycetes",
                                         "Tremellomycetes",
                                         "Exobasidiomycetes",
                                         "Microbotryomycetes"),
                             seq %in% longtree$tip.label)

basidtree <- phylobase::phylo4(longtree) %>%
  phylobase::subset(mrca = basids$seq) %>%
  as("phylo")

basids2 <- treetax %>% filter(seq %in% basidtree$tip.label)

ggtree(groupOTU(basidtree, split(basids2$seq, basids2$Order)), layout = "circular") +
  aes(color = group) +
  scale_color_brewer(type = "qual", palette = "Paired", limits = unique(na.omit(basids2$Order))) +
  theme(legend.position = NULL)
@
\end{frame}

\begin{frame}{Correlogram}
<<correlogram, fig.width = 12>>=
bind_rows(as.data.frame(bc_corr$mantel.res) %>%
            mutate(tdist = "same", metric = "bray"),
          as.data.frame(bc_corr2$mantel.res) %>%
            mutate(tdist = "diff", metric = "bray"),
          as.data.frame(unif_corr$mantel.res) %>%
            mutate(tdist = "same", metric = "unifrac"),
          as.data.frame(unif_corr2$mantel.res) %>%
            mutate(tdist = "diff", metric = "unifrac")) %>%
  mutate(Significant = `Pr(corrected)` < 0.05) %>%
  ggplot(aes(class.index, Mantel.cor, group = tdist, color = tdist, shape = Significant)) +
  geom_point(size = 3) +
  geom_line(alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
  # geom_line(aes(y = exp(predict(bc_fit)))) +
  scale_y_continuous(limits = c(-0.05, NA), name = "Spatial Distance correlation") +
  scale_color_discrete(name = "Year") +
  xlab("distance (m)") +
  facet_wrap("metric") +
  scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 1))
@
\end{frame}

\end{document}