### 2019-10-09 writing/LSUx.Rmd
LSUx and reconstruct.  Intended to run up to consensus.

---
title: Comparison of DADA2 performance on LSUx domains (including ITS and 5.8S)
  rDNA
author: "Brendan Furneaux"
date: "October 2, 2019"
output:
  beamer_presentation:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, device = "tikz")
library(ggplot2)
theme_set(theme_bw(base_size = 24) + theme(legend.direction = "vertical"))
library(drake)
library(magrittr)
targets <- drake_cache()$list()
expand.grid(plate = c("001", "002"),
            region = c("ITS1", "5_8S", "ITS2", "LSU1", "D1", "LSU2", "D2",
                       "LSU3", "D3", "LSU4", "ITS", "LSU", "32S", "long"))

table <- tibble::tibble(name = targets[startsWith(targets, "nochim_pb_500")]) %>%
  tidyr::extract(col = name, into = c("plate", "region"),
                 regex = "nochim_pb_500_(00[12])_(.+)",
                 remove = FALSE) %>%
  dplyr::mutate_at("region", factor, levels = c("ITS1", "5_8S", "ITS2", "LSU1",
                                                "D1", "LSU2", "D2", "LSU3",
                                                "D3", "LSU4", "ITS", "LSU",
                                                "32S", "long")) %>%
  dplyr::group_by(region) %>%
  dplyr::mutate(nochim = lapply(name, readd, character_only = TRUE)) %>%
  dplyr::summarize(nochim = list(dada2::mergeSequenceTables(tables = nochim)),
                   reads = sum(nochim),
                   ASVs = ncol(nochim),
                   length_min = min(nchar(colnames(nochim))),
                   length_q1 = quantile(rep(nchar(colnames(nochim)),
                                            colSums(nochim)), 0.25),
                   length_med = median(rep(nchar(colnames(nochim)),
                                           colSums(nochim))),
                   length_mean = mean(rep(nchar(colnames(nochim)),
                                          colSums(nochim))),
                   length_q3 = quantile(rep(nchar(colnames(nochim)),
                                            colSums(nochim)), 0.75),
                   length_max = max(nchar(colnames(nochim))))
```

## Reads vs. ASVs

```{r}
ggplot(table, aes(reads, ASVs, color = length_med, label = region)) +
  geom_text() +
  scale_color_gradientn(colors = c("blue", "red", "green1"), limits = c(30, 2000), breaks = c(50, 150, 400, 1000), trans = "log10", name = "median length") +
  xlab("mapped reads")
```

## Length distributions
```{r}
ggplot(table, aes(x = length_med, y = reads, label = region)) +
  geom_text() +
  xlab("length (bp)")
```

## Length distributions - Variable regions
```{r}
lengths_table <- table %>%
  mutate(
    nreads = map(nochim, colSums),
    length = map(nochim, ~ nchar(colnames(.)))
  ) %>%
  select(region, length, nreads) %>%
  unnest() %>%
  group_by(region, length) %>%
  summarize(nreads = sum(nreads))

lengths_table %>%
  filter(region %in% c("ITS1", "ITS2", "D1", "D2", "D3")) %>%
  dplyr::slice(rep(1:nrow(.), .$nreads)) %>%
  ggplot(aes(length)) +
  geom_density(bw = 2) +
  facet_wrap(~region,ncol = 2, strip.position = "right") +
  scale_y_continuous(breaks = NULL, labels = NULL)
```

## Length distributions -- conserved regions

```{r}
lengths_table %>%
  filter(region %in% c("5_8S", "LSU1", "LSU2", "LSU3", "LSU4")) %>%
  dplyr::slice(rep(1:nrow(.), .$nreads)) %>%
  ggplot(aes(length)) +
  geom_density(bw = 1) +
  facet_wrap(~region,ncol = 2, strip.position = "right") +
  scale_y_continuous(breaks = NULL, labels = NULL) +
  scale_x_continuous(limits = c(0, 300))
```

## Length distributions -- combined regions

```{r}
lengths_table %>%
  filter(region %in% c("ITS", "LSU", "32S", "long")) %>%
  dplyr::slice(rep(1:nrow(.), .$nreads)) %>%
  ggplot(aes(length)) +
  geom_density(bw = 10) +
  facet_wrap(~region,ncol = 2, strip.position = "right") +
  scale_y_continuous(breaks = NULL, labels = NULL) +
  scale_x_continuous(limits = c(0, 2000))
```